Language Skills in an Ethnically Segmented Labour market: Estonia 1989 -- 2012

### 2018-07-10 paper/ijm.tex
Added table and figure to section 7

The most surprising result in our analysis so far--the fact that
Russian men earn virtually no Estonian language premium while women do--warrants a
deeper look at the data.  We compute the average wage for the Russian male wage-earners
along the three
dimensions: Estonian language skills, residence in Tallinn metro area,
and education.  The metropolitan area is one of the two major regions
where the Russian-speaking population is concentrated, and it's
booming economy may be very different from that in the North-East.
Even more, while Russians form a large minority in and around Tallinn,
they are in majority in the North-East.  
In addition, to correct very strong wage and language skill trends through these
years, we also present the average
residuals from an OLS regression that explains log wage by year fixed
effects (Table~\ref{tab:loc-lang-edu}). 
\fxwarning{ot: Sven, can you redo this table?  Use exactly the same
  sample as you used.  Would also be good to put both et1 and et0 on
  the same row}
\fxwarning{SKB: Will try. Although I still do not understand, what the residual is telling me and why the wages are so high. Are the wages in EEK? What does ET stand for? The table needs more explanation or at least a proper legend.}
The table shows that the picture is in fact quite complex.  We can see
that there is little to no gain from Estonian skills for the
largest group, high-school educated workers.  Even more, outside of
the metropolitan area, the lowest and highest education groups
have in fact negative language premium.  However, these two groups
enjoy quite a noticeable premium in the Tallinn region.  The
opposite-signed effects in for these two education groups suggests
that a
selective internal migration may play a role.

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Fri Jul  6 15:59:47 2018
\begin{table}[ht]
  \centering
  \caption{Average wage and time-trend corrected residual across
    location and education groups for Russian men}
  \begin{tabular}{lrlrrr}
    \toprule
    Tallinn & education & ET & wage & residual & \# obs \\ 
    \midrule
    Tallinn & $<$ HS & 0 & 4313.38 & -0.19 & 255 \\ 
    Tallinn & $<$ HS & 1 & 5351.11 & -0.05 & 219 \\ 
    Tallinn & HS     & 0 & 5803.21 & 0.05 & 1292 \\ 
    Tallinn & HS      & 1 & 5666.77 & 0.05 & 1296 \\ 
    Tallinn & college & 0 & 6562.57 & 0.16 & 212 \\ 
    Tallinn & college & 1 & 8453.14 & 0.30 & 518 \\ 
    other   & $<$ HS & 0 & 3877.84 & -0.13 & 461 \\ 
    other   & $<$ HS & 1 & 4496.62 & -0.18 & 230 \\ 
    other   & HS     & 0 & 4838.65 & -0.09 & 2233 \\ 
    other   & HS      & 1 & 5351.51 & -0.05 & 960 \\ 
    other   & college & 0 & 6283.38 & 0.17 & 278 \\ 
    other   & college & 1 & 6072.06 & 0.13 & 263 \\ 
    \bottomrule
  \end{tabular}
  \label{tab:loc-lang-edu}
\end{table}

We can conclude the table~\ref{tab:loc-lang-edu} that our
most suprising finding, no positive effect of Estonian language
skills for Russian men, is primarily driven by workers with
high-school degree.  It also suggests that highly educated workers in
the metropolitan area enjoy, in fact, a substantial language premium.
However, this group is not very large.

>>>>

The most surprising result in our analysis so far--the fact that
Russian men earn virtually no Estonian language premium while women do--warrants a
deeper look at the data.  We compute the average wage for the Russian male wage-earners
along the three
dimensions: Estonian language skills, residence in Tallinn metro area,
and education.  The metropolitan area is one of the two major regions
where the Russian-speaking population is concentrated, and it's
booming economy may be very different from that in the North-East.
Even more, while Russians form a large minority in and around Tallinn,
they are in majority in the North-East.  
In addition, to correct very strong wage and language skill trends through these
years, we also present the average
residuals from an OLS regression that explains log wage by year fixed
effects (Table~\ref{tab:loc-lang-edu}). 
\fxwarning{ot: Sven, can you redo this table?  Use exactly the same
  sample as you used.  Would also be good to put both et1 and et0 on
  the same row}
\fxwarning{SKB: Will try. Although I still do not understand, what the residual is telling me and why the wages are so high. Are the wages in EEK? What does ET stand for? The table needs more explanation or at least a proper legend.}

The table shows that the picture is in fact quite complex.  We can see
that there is little to no gain from Estonian skills for the
largest group, high-school educated workers.  Even more, outside of
the metropolitan area, the lowest and highest education groups
have in fact negative language premium.  However, these two groups
enjoy quite a noticeable premium in the Tallinn region.  The
opposite-signed effects in for these two education groups suggests
that a
selective internal migration may play a role.

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Fri Jul  6 15:59:47 2018
\begin{table}[ht]
  \centering
  \caption{Average wage and time-trend corrected residual across
    location and education groups for Russian men}
  \begin{tabular}{lrlrrr}
    \toprule
    Tallinn & education & ET & wage & residual & \# obs \\ 
    \midrule
    Tallinn & $<$ HS & 0 & 4313.38 & -0.19 & 255 \\ 
    Tallinn & $<$ HS & 1 & 5351.11 & -0.05 & 219 \\ 
    Tallinn & HS     & 0 & 5803.21 & 0.05 & 1292 \\ 
    Tallinn & HS      & 1 & 5666.77 & 0.05 & 1296 \\ 
    Tallinn & college & 0 & 6562.57 & 0.16 & 212 \\ 
    Tallinn & college & 1 & 8453.14 & 0.30 & 518 \\ 
    other   & $<$ HS & 0 & 3877.84 & -0.13 & 461 \\ 
    other   & $<$ HS & 1 & 4496.62 & -0.18 & 230 \\ 
    other   & HS     & 0 & 4838.65 & -0.09 & 2233 \\ 
    other   & HS      & 1 & 5351.51 & -0.05 & 960 \\ 
    other   & college & 0 & 6283.38 & 0.17 & 278 \\ 
    other   & college & 1 & 6072.06 & 0.13 & 263 \\ 
    \bottomrule
  \end{tabular}
  \label{tab:loc-lang-edu}
\end{table}
\fxwarning{SKB: Copied and modified table above. }
\begin{table}[ht]
	\centering
	\caption{Average wage and time-trend corrected language parameter across
		location and education groups for Russian men}
	\begin{tabular}{lrlrrr}
		\toprule
		Tallinn & education & ET & wage & parameter & \# obs \\ 
		\midrule
		Tallinn & $<$ HS & 0 & 334 & -0.16\onepc & 255 \\ 
		Tallinn & $<$ HS & 1 & 424 & -0.08 & 219 \\ 
%		Tallinn & HS     & 0 &476 & 0.05 & 1292 \\ 
		Tallinn & HS      & 1 & 512 & -0.05 \fivepc & 1296 \\ 
		Tallinn & college & 0 & 520 & -0.02 & 212 \\ 
		Tallinn & college & 1 & 737 & 0.05 & 518 \\ 
		Ida-Viru & $<$ HS & 0 & 396 &  -0.20\onepc & 379 \\ 
		Ida-Viru & $<$ HS & 1 & 269 & -0.23\onepc & 20 \\ 
		Ida-Viru & HS     & 0 & 337 & -0.23\onepc & 1889 \\ 
		Ida-Viru & HS      & 1 & 368 & -0.21\onepc & 252 \\ 
		Ida-Viru & college & 0 & 451 & -0.07 & 246 \\ 
		Ida-Viru & college & 1 & 493 & -0.06 & 154 \\ 
		other   & $<$ HS & 0 & 279 & -0.23\onepc & 83 \\ 
		other   & $<$ HS & 1 & 334 & -0.19\onepc & 212 \\ 
		other   & HS     & 0 & 378& -0.13\onepc & 353 \\ 
		other   & HS      & 1 & 389 & -0.09\onepc & 713\\ 
		other   & college & 0 & 788 & 0.03 & 32\\ 
		other   & college & 1 & 414 & -0.11 & 110 \\ 
		\bottomrule
	\end{tabular}
	\label{tab:loc-lang-edu_m3}
\end{table}

\fxwarning{SKB: Added graph showing the same parameter estimates. Maybe a better formatted graph like this shows the story better than a table}
\begin{figure}[hbtp]
	\centering
	\caption{Coefficient plot of Model 3 for Russian Men separated by Region and Education}
	\label{fig:russianmenlanguagecoefsm3}
	\includegraphics[width=1.0\linewidth]{russian_men_language_coefs_m3}
\end{figure}

We can conclude the table~\ref{tab:loc-lang-edu} that our
most surprising finding, no positive effect of Estonian language
skills for Russian men, is primarily driven by workers with
high-school degree.  It also suggests that highly educated workers in
the metropolitan area enjoy, in fact, a substantial language premium.
However, this group is not very large.
